Thesis outline

### 2017-06-06 KellyDouglass/Thesis/Thesis_sections/GV.tex
Small-scale intro, GV edits

\chapter[Green valley galaxies]{Green valley galaxies}

@@

% Feedback from AGN (strongly affects massive galaxies)
% Feedback from SN (dominates feedback in dwarfs)
% gas stripping interactions
% tidally-triggered star formation
There are a number of different astrophysical processes which simultaneously 
evolve a galaxy's stellar and gas content.  AGN feedback strongly affects 
massive galaxies, while dwarf galaxy feedback is dominated by supernovae.  Both 
AGN and supernovae can inject enough energy into a galaxy's ISM to prevent its 
cooling, thereby quenching star formation.  In addition to internal mechanisms, 
galaxies will often experience interactions in more dense regions.  Depending on 
the properties of the galaxies involved (mass, speed, etc.), these interactions 
can either strip gas from the galaxy (removing a primary source of fuel for 
future star formation), or they can trigger a burst of star formation.

@@

The SDSS Data Release 7 \citep[DR7]{Abazajian09} is a wide-field multi-band 
imaging and spectroscopic survey.  It uses drift scanning to map approximately 
one-quarter of the northern sky.  A dedicated 2.5-meter telescope at the Apache 
Point Observatory in New Mexico takes photometric data in the five-band SDSS 
system --- $u$, $g$, $r$, $i$, and $z$ \citep{Fukugita96,Gunn98}.  Galaxies with 
a Petrosian $r$-band apparent magnitude $m_r < 17.77$ are selected for 
spectroscopic analysis \citep{Lupton01,Strauss02}.  Gas-phase chemical 
abundances are calculated via the Direct $T_e$ method using emission-line fluxes 
measured by the MPA-JHU catalog\footnote{Available at 
\url{http://www.mpa-garching.mpg.de/SDSS/DR7}} and published in Douglass \& 
Vogeley (2017, in prep).  The large-scale environment of the galaxies is based 
on the void catalog compiled by \cite{Pan12}.  This void catalog is constructed 
from SDSS DR7 using the VoidFinder algorithm of \cite{Hoyle02}.  Galaxies which 
are located within the identified voids are labeled as void galaxies; those 
which do not reside in a void are designated as wall galaxies.  Because the 
minimum diameter of a void is defined as 10 \hMpc , the large-scale environment 
of galaxies within 5 \hMpc of the edge of the SDSS DR7 footprint cannot be 
described, so their environment is labeled as unknown.

@@

If the green valley galaxies represent those galaxies transitioning between the 
blue sequence and red cloud, then their stellar masses should overlap the masses 
in both these populations.  The distribution of the stellar masses for the 
galaxies in SDSS DR7 is shown in Fig. \ref{fig:M_hist}.  We can see that the 
galaxies in the green valley have stellar masses comparable to those in the red 
cloud.  The stellar masses of those galaxies in the blue sequence range from the 
lowest mass objects up through the stellar masses of the green valley galaxies.  
The distribution of stellar masses of galaxies in the blue sequence should be 
much broader than the range of masses seen in the red cloud galaxies, since 
those in the blue sequence are forming stars and slowly building up their mass.  
Galaxies in the red cloud are fairly stagnant in their star formation, so the 
range of stellar masses of these galaxies will be more narrow than that of the 
blue sequence.  If the green valley galaxies are transitioning between the two 
populations due to a quenching of star formation or a sudden burst of star 
formation, then we would expect their mass range to cover the high-mass end of 
the blue sequence and most of the red cloud.  Indeed, this is what we see in 
Fig. \ref{fig:M_hist}.

@@

Again, if green valley galaxies are transitioning between the blue sequence and 
red cloud, then they should have intermediate (s)SFRs.  The distribution of SFR 
and sSFR for the galaxies in SDSS DR7 are shown in Fig. \ref{fig:SFR_hist}, with 
the galaxies separated by their morphological types.  We see that the 
distribution of SFRs for galaxies in the green valley peaks around 
$\log(\text{SFR}) \sim -0.5$, in between the ranges occupied by the red cloud 
and blue sequence galaxies.  A similar distribution is shown for the sSFRs, 
although the green valley galaxies extend well into the range of sSFRs occupied 
by the galaxies in the red cloud.  \cite{Salim14a} describes transitional 
galaxies (those in the green valley) to have lower sSFRs than galaxies in the 
blue sequence, specifically $-11.8 < \log(\text{sSFR}) < -10.8$.  While our 
results do indicate that the green valley galaxies have lower sSFRs than those 
in the blue sequence, we find that their sSFRs extend well into the red cloud 
range, or to $\log(\text{sSFR}) \sim -12$.

@@

The stellar mass range of galaxies in the green valley in Fig. \ref{fig:M_hist} 
shows us that these galaxies have completed most of their star formation.  They 
have acquired enough stars to have converted most of their gas into heavier 
elements.  Combined with the intermediate SFRs and intermediate-to-low sSFRs 
seen in Fig. \ref{fig:SFR_hist}, this tells us that these galaxies are no longer 
forming stars at a high rate.  This is most likely due to an extinction of 
available cool gas from which to form stars.  These galaxies are reaching a 
natural point in their evolution where they transition from an active star 
formation period to a quieter, more quiescent lifestyle.

@@

Fig. \ref{fig:Mr_bin} shows that a higher fraction of faint wall galaxies are in 
the green valley than faint void galaxies.  This would indicate that void 
galaxies are slightly behind wall galaxies in their evolution, matching 
predictions based on the $\Lambda$CDM cosmology.  These results also coincide 
with the observations of \cite{Douglass17b} and Douglass \& Vogeley (2017, in 
prep).  They find that, for dwarf galaxies ($M_r > -17$), those which exist in 
voids may have commenced their star formation at a later time and have not 
experienced the same cosmic downsizing as dwarf galaxies residing in denser 
regions.  The higher fraction of wall dwarf galaxies in the green valley might 
also point to an environmental influence on the reason for transitioning through 
the green valley, such as ram pressure stripping, which would occur much less 
frequently in the void environment.  However, the existence of void galaxies in 
the green valley indicate that galaxy interactions cannot be the only quenching 
mechanism.

@@

During my estimation of the chemical content of galaxies with $M_r > -20$, I 
note that the oxygen abundances of galaxies classified as AGN by their 
emission line ratios in the BPT diagrams exhibit higher nitrogen abundances 
(but normal oxygen abundance values) when compared to star-forming galaxies of 
a similar stellar mass and large-scale environment.  The same abundance 
signature is seen in the the oxygen and nitrogen abundance results seen in Fig. 
\ref{fig:Z_hist} of the green valley galaxies.  This might indicate a 
link between green valley galaxies and a host AGN.  This would support other 
studies of the transition of galaxies from the blue sequence to the red cloud, 
which attribute the quenching of star formation to accretion of a central black 
hole \citep[e.g.,][]{Croton06,Stasinska08}.

@@

Based on our analysis, green valley galaxies have stellar masses comparable to 
those in the red cloud.  They have intermediate SFRs, and low-to-intermediate 
sSFRs.  Green valley galaxies also have high gas-phase nitrogen abundances (N/H), 
resulting in high N/O ratios.  While their SFRs show that something has quenched 
their star formation, their stellar masses inform us that this is not due to any 
premature quenching mechanism.  The high nitrogen abundances in the green valley 
galaxies indicates that the galaxies are either no longer forming stars (since 
nitrogen is produced in lower mass stars than oxygen), or that the galaxies were 
able to produce both primary and secondary nitrogen (heavy elements were present 
during the last few star formation episodes to permit the CNO cycle to commence 
earlier).  This chemical abundance pattern (normal oxygen, high nitrogen, and 
high N/O ratio) is also seen in galaxies classified as AGN, indicating that the 
galaxies in the green valley have an AGN.

There is a higher fraction of faint wall galaxies in the green valley than in 
the void regions.  In conjunction with the conclusions of \cite{Douglass17b} and 
Douglass \& Vogeley (2017, in prep), this indicates that void dwarf galaxies are 
less evolved than dwarf galaxies in denser environments.  It would be beneficial 
to investigate the small-scale ($\sim 1$ \hMpc) environment of galaxies in the 
green valley, to determine if any interactions are responsible for the 
transitional state.

>>>>

\chapter[Green valley galaxies]{Identifying green valley galaxies}

@@

% Feedback from AGN (strongly affects massive galaxies)
% Feedback from SN (dominates feedback in dwarfs)
% gas stripping interactions
% tidally-triggered star formation
There are a number of different astrophysical processes which simultaneously 
evolve a galaxy's stellar and gas content.  AGN feedback strongly affects 
massive galaxies, while dwarf galaxy feedback is dominated by supernovae.  Both 
AGN and supernovae can inject enough energy into a galaxy's ISM to prevent its 
cooling, thereby quenching star formation.  In addition to internal mechanisms, 
galaxies will often experience interactions in denser environmental regions.  
Depending on the properties of the galaxies involved (mass, speed, etc.), these 
interactions can either strip gas from the galaxy (removing a primary source of 
fuel for future star formation), or they can trigger a burst of star formation.

@@

The SDSS Data Release 7 \citep[DR7]{Abazajian09} is a wide-field multi-band 
imaging and spectroscopic survey that uses drift scanning to map approximately 
one-quarter of the northern sky.  A dedicated 2.5-meter telescope at the Apache 
Point Observatory in New Mexico takes photometric data in the five-band SDSS 
system --- $u$, $g$, $r$, $i$, and $z$ \citep{Fukugita96,Gunn98}.  Galaxies with 
a Petrosian $r$-band apparent magnitude $m_r < 17.77$ are selected for 
spectroscopic analysis \citep{Lupton01,Strauss02}.  Gas-phase chemical 
abundances are calculated via the Direct $T_e$ method using emission-line fluxes 
measured by the MPA-JHU catalog\footnote{Available at 
\url{http://www.mpa-garching.mpg.de/SDSS/DR7}} and published in Douglass \& 
Vogeley (2017, in prep).  The large-scale environment of the galaxies is based 
on the void catalog compiled by \cite{Pan12}.  This void catalog is constructed 
from SDSS DR7 using the VoidFinder algorithm of \cite{Hoyle02}.  Galaxies which 
are located within the identified voids are labeled as void galaxies; those 
which do not reside in a void are designated as wall galaxies.  Because the 
minimum diameter of a void is defined as 10 \hMpc , the large-scale environment 
of galaxies within 5 \hMpc of the edge of the SDSS DR7 footprint cannot be 
described, so their environment is labeled as unknown.

@@

If the green valley galaxies represent those galaxies transitioning between the 
blue sequence and red cloud, then their stellar masses should overlap the masses 
in both these populations.  The distribution of the stellar masses for the 
galaxies in SDSS DR7 is shown in Fig. \ref{fig:M_hist}.  We can see that the 
galaxies in the green valley have stellar masses comparable to those in the red 
cloud.  The stellar masses of those galaxies in the blue sequence range from the 
lowest mass objects up through the stellar masses of the green valley galaxies.  
The distribution of stellar masses of blue sequence galaxies should be much 
broader than the range of masses seen in the red cloud galaxies, since those in 
the blue sequence are slowly building up their stellar mass by forming stars 
from both their gas reservoirs and cool gas falling onto the galaxy.  Galaxies 
in the red cloud are fairly stagnant in their star formation, so the range of 
stellar masses of these galaxies will be more narrow than that of the blue 
sequence.  If the green valley galaxies are transitioning between the two 
populations due to a quenching of star formation or a sudden burst of star 
formation, then we would expect their mass range to cover the high-mass end of 
the blue sequence and most of the red cloud.  These expectations are met in Fig. 
\ref{fig:M_hist}.

@@

If green valley galaxies are transitioning between the blue sequence and red 
cloud, then they should have intermediate (s)SFRs.  The distribution of SFR and 
sSFR for the galaxies in SDSS DR7 are shown in Fig. \ref{fig:SFR_hist}, with the 
galaxies separated by their morphological types.  We see that the distribution 
of SFRs for galaxies in the green valley peaks around 
$\log(\text{SFR}) \sim -0.5$, in between the ranges occupied by the red cloud 
and blue sequence galaxies.  A similar distribution is shown for the sSFRs, 
although the green valley galaxies extend well into the range of sSFRs occupied 
by the galaxies in the red cloud.  \cite{Salim14a} describes transitional 
galaxies (those in the green valley) to have lower sSFRs than galaxies in the 
blue sequence, specifically $-11.8 < \log(\text{sSFR}) < -10.8$.  While our 
results do indicate that green valley galaxies have lower sSFRs than those in 
the blue sequence, we find that their range of sSFR is almost double the width 
of that described in \cite{Salim14a}.  The green valley galaxies have a sSFR 
within the range $-12 \lesssim \log(\text{sSFR}) \lesssim -10$.

@@

The range of stellar masses in galaxies from the green valley in Fig. 
\ref{fig:M_hist} shows us that these galaxies have completed most of their star 
formation.  They have acquired enough stars to have converted much of their gas 
into heavier elements.  Combined with the intermediate SFRs and 
intermediate-to-low sSFRs seen in Fig. \ref{fig:SFR_hist}, this tells us that 
these galaxies are no longer forming stars at a high rate.  This is most likely 
due to an extinction of available cool gas from which to form stars.  These 
galaxies are reaching a natural point in their evolution where they transition 
from an active star formation period to a quieter, more quiescent lifestyle.

@@

Fig. \ref{fig:Mr_bin} shows a higher fraction of faint wall galaxies than faint 
void galaxies are in the green valley.  This would indicate that void galaxies 
are slightly behind wall galaxies in their evolution, matching predictions based 
on the $\Lambda$CDM cosmology.  These results also coincide with the 
observations of \cite{Douglass17b} and Douglass \& Vogeley (2017, in prep).  
They find that, for dwarf galaxies ($M_r > -17$), those which exist in voids may 
have commenced their star formation at a later time and have not experienced the 
same cosmic downsizing as dwarf galaxies residing in denser regions.  The higher 
fraction of wall dwarf galaxies in the green valley might also point to an 
environmental influence on the reason for transitioning through the green 
valley, such as ram pressure stripping, which would occur much less frequently 
in the void environment.  However, the existence of void galaxies in the green 
valley indicate that galaxy interactions cannot be the only star formation 
quenching mechanism.

@@

During my estimation of the chemical content of galaxies with $M_r > -20$, I 
note that the oxygen abundances of galaxies classified as having an AGN by their 
emission line ratios in the BPT diagrams exhibit higher nitrogen abundances and 
normal oxygen abundance values when compared to star-forming galaxies of a 
similar stellar mass and large-scale environment.  The same abundance signature 
is seen in the the oxygen and nitrogen abundance results seen in Fig. 
\ref{fig:Z_hist} of the green valley galaxies.  This might indicate a 
link between green valley galaxies and a host AGN.  This would support other 
studies of the transition of galaxies from the blue sequence to the red cloud, 
which attribute the quenching of star formation to accretion of a central black 
hole \citep[e.g.,][]{Croton06,Stasinska08}.

@@

Based on our analysis, green valley galaxies have stellar masses comparable to 
those in the red cloud.  They have intermediate SFRs, and low-to-intermediate 
sSFRs.  Green valley galaxies also have high gas-phase nitrogen abundances (N/H), 
resulting in high N/O ratios.  While their SFRs show that something has quenched 
their star formation, their stellar masses inform us that this is not due to any 
premature quenching mechanism.  The high nitrogen abundances in the green valley 
galaxies indicate that the galaxies are either no longer forming stars (since 
nitrogen is produced in lower mass stars than oxygen), or that the galaxies were 
able to produce both primary and secondary nitrogen (heavy elements were present 
during the last few star formation episodes to permit the CNO cycle to commence 
earlier).  This chemical abundance pattern (normal oxygen, high nitrogen, and 
high N/O ratio) is also seen in galaxies classified as AGN, indicating that the 
galaxies in the green valley have an AGN.

There is a higher fraction of faint wall galaxies than faint void galaxies in 
the green valley.  In conjunction with the conclusions of \cite{Douglass17b} and 
Douglass \& Vogeley (2017, in prep), this indicates that void dwarf galaxies are 
less evolved than dwarf galaxies in denser environments.  It would be beneficial 
to investigate the small-scale ($\sim 1$ \hMpc) environment of galaxies in the 
green valley, to determine if any interactions are responsible for the 
transitional state.

### 2017-06-06 KellyDouglass/Thesis/Thesis_sections/smallScaleEnvironment.tex
Small scale theory update
Small-scale intro, GV edits

Morphology-density relation \cite{Dressler80}

@@

We can also calculate the distance between two galaxies by ignoring the effects 
of their relative velocities.

>>>>

% Large-scale environment influence
Large galaxy redshift surveys have shown that the large-scale structure of the 
distribution of galaxies in the universe is similar to that of a 
three-dimensional cosmic web \citep{Bond96}, where galaxy clusters are connected 
by thin filaments of galaxies and are separated by voids (large, underdense 
regions that occupy close to 60\% of space).  Over the last few decades, the 
Sloan Digital Sky Survey \citep{Abazajian09,Ahn12} have propelled the research 
into the influence of the large-scale environment on the formation and evolution 
of galaxies.

Because gravitational clustering within a void proceeds as if in a very 
low-density universe, cosmic voids are an important environment for studying 
galaxy formation \citep[see][for a review]{vandeWeygaert11}.  The $\Lambda$CDM 
cosmology predicts that galaxies which formed in voids to have lower masses and 
be retarded in their star formation when compared to those in more dense 
environments \citep{Gottlober03,Goldberg05,Cen11}.  The effects of the void 
environment should be most obvious in dwarf galaxies due to their minimal 
gravitational potential.  As a result, they are more sensitive to astrophysical 
effects such as cosmological reionization, internal feedback from supernova and 
photoheating from star formation, small-scale details of dark matter halo 
assembly, and the properties of dark matter.

Observations have shown that the properties of dwarf galaxies vary dramatically 
with the environment \citep[e.g.,][]{Ann08,Geha12}.  Void galaxies have been 
found to have lower stellar mass \citep{Hoyle05,Croton05,Moorman15}, be bluer 
and of a later type \citep{Grogin00,Rojas04,Patiri06,Park07,vonBendaBeckmann08,
Hoyle12}, have higher star formation rates \citep{Rojas05,Moorman15,Beygu16}, 
and be more gas rich \citep{Kreckel12,Moorman16,Jones16} than galaxies in denser 
regions.

% Small-scale environment - what has been done so far?
In conjunction with the large-scale environment, the small-scale environment 
($\sim 1$ \hMpc) has also been found to influence a galaxy's evolution.  There 
exits a well-established morphology-density relation \cite{Dressler80}, where 
the fraction of late-type galaxies is inversely proportional to the local 
density.  \cite{Ellison09} determined that a galaxy's local environment 
influences a galaxy's evolution more than its large-scale environment.  
Likewise, \cite{Rupke08} concluded that interacting galaxies have suppressed 
metallicities, because interactions induce flows of hydrogen.  \cite{Park09} 
found that, for galaxies with $M_r < -19$, galaxy interactions out to the virial 
radius of the nearest neighbor influence the evolution of the target galaxy.  
They determine that the large-scale environment has a minimal effect on the 
evolution of a galaxy once the luminosity and morphology are taken into account.

We want to understand if the small-scale environment is more influential in a 
galaxy's evolution than its large-scale environment.  Additionally, within the 
small-scale environment, are tidal influences (due to the nearest neighbor 
galaxy) or gravitational potentials (from the nearest galaxy group) more 
influential in a galaxy's evolution?  A group consists of multiple galaxies that 
share the same dark matter halo.  A group member will gravitationally interact 
with all other group members, not just its nearest neighbor.  And a galaxy near 
a group, but not within it, may also experience strong tidal effects from the 
group similar to (but stronger than) those from a neighboring galaxy.

@@

We can also calculate the distance between two galaxies by ignoring the effects 
of their relative velocities.  Instead of projecting both galaxies onto a sphere 
with radius $\bar{r}$, we instead project each onto its own sphere of radius 
\begin{equation}
    r = \frac{cz}{H_0}
\end{equation}
We then convert to Cartesian coordinates and use the three-dimensional 
Pythagorean theorem to calculate the distance between the two galaxies.
